---
title: "PSL quiz 2"
author: "Brendan Shea"
date: '2022-09-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
Q1-Q3 are related to the  Auto dataset, which is attached below.

```{r}
quiz2_data <- read.csv("quiz2_data.data")

head(quiz2_data)
```



Using  mpg as response and  displacement as predictor, perform a simple linear regression (round to the  2nd decimal place ). The slope is ______.

```{r}
fit_disp = lm(mpg ~ displacement, data =  quiz2_data)
fit_disp
round(fit_disp$coefficients[2],2)
```
## Question 2
Continue with Question 1. The R-square is ______ (the answer should be a number between 0 and 1; round to the  2nd decimal place ).

```{r}
round(
    summary(fit_disp)$r.squared, 2)
```

## Question 3
Fit a multiple linear regression model using  mpg as response and the other seven variables except "name" (i.e., cylinders, displacement, horsepower, weight, acceleration, year, origin ) as predictors.

 
The coefficient associated with  displacement is ______ (round to the 2nd decimal place ).
```{r}
fit3 = lm(mpg ~ . - name, data = quiz2_data)
fit3

```

```{r}
round(fit3$coefficients[3],2)
```

## Question 4

Question 4
Q4-Q21 are related to a linear regression model that predicts starting salary after graduation (in thousands of dollars). The model has five predictors,

$$X_1$$ = GPA,

$$X_2$$ = IQ,

$$X_3$$ = Gender (1 for Female and 0 for Male),

$$X_4$$ = Interaction between GPA and IQ (i.e., GPA$$*$$IQ), and

$$X_5$$ = Interaction between GPA and Gender (i.e., GPA$$*$$Gender).

Suppose we use least squares to fit the model, and the intercept $$b_0$$ and the five coefficients (associated with $$X_1$$ to $$X_5$$) are:

$$b_0 = 50; b_1 =20, b_2 =0.07, b_3 =35, b_4 = 0.01, b_5 =-10$$


```{r}
b_0_int = 50
b_1_gpa =20
b_2_iq =0.07
b_3_gender = 35
b_4_gpa_v_iq = 0.01
b_5_gpa_v_gender =-10
```


Predict the salary (in thousands of dollars, e.g., if salary is 10,000, your input should be 10) of a female with IQ of 100 and a GPA of 4.0.

```{r}
x_1 = 4.0
x_2 = 100
x_3 = 1

yhat_4 = b_0_int + b_1_gpa * x_1 + b_2_iq * x_2 + b_3_gender * x_3 + b_4_gpa_v_iq * x_1 * x_2 + b_5_gpa_v_gender* x_1 * x_3
yhat_4
```

## Question 5
Which  two of the following are correct?

```{r}

for(x_1 in seq(from = 3, to = 4, by = 0.1)){
    for(x_3 in 0:1){
    yhat_5 = b_0_int + b_1_gpa * x_1 + b_2_iq * x_2 + b_3_gender * x_3 + b_4_gpa_v_iq * x_1 * x_2 + b_5_gpa_v_gender* x_1 * x_3
    cat("gender =", x_3, "gpa = ", x_1, yhat_5, "\n")
    }
}
```


1 point



X For a fixed value of IQ and GPA, males earn more on average than females provided that the GPA is higher than 3.5.


X For a fixed value of IQ and GPA, females earn more on average than males provided that the GPA is less than 3.5.

## Question 6
Suppose we record the IQ as relative IQ to a base score 100. That is, the new $$X_2$$ becomes the original $$X_2/100$$: IQ score 100 is now recorded as 1, and IQ score 150 is now recorded as 1.5.

 
If we use the same data set (everything is the same except the $$X_2$$ and $$X_4$$ columns are now scaled by 1/100) to fit a linear regression model, which statement about the estimated LS coefficients is correct?

```{r}
x_1 = 4.0
x_2 = 1
x_3 = 1

yhat_4 = b_0_int + b_1_gpa * x_1 + b_2_iq * x_2 * 100 + b_3_gender * x_3 + b_4_gpa_v_iq * x_1 * x_2 * 100 + b_5_gpa_v_gender* x_1 * x_3
yhat_4
```


The new estimates for $$b_2$$ and $$b_4$$ are 7 and 1, respectively; all other coefficients stay the same.

## Questions 7, 8, 9, 10
Based on the LS coefficients given in  Question 4, given IQ and GPA, the prediction for the salary of a  female student can be written as the following linear combination

$$
w_0 + w_1*\text{GPA} + w_2*\text{IQ} + w_3*(\text{GPA}*\text{IQ})
$$

```{r}
yhat_7 = (b_0_int + b_3_gender) + (b_1_gpa + b_5_gpa_v_gender) * x_1 + (b_2_iq * x_2)  + (b_4_gpa_v_iq * x_1 * x_2)
yhat_7
```

```{r}
(w0 = b_0_int + b_3_gender)
(w1 = b_1_gpa + b_5_gpa_v_gender)
(w2 = b_2_iq)
(w3 = b_4_gpa_v_iq)
w0 + w1 * x_1 + w2 * x_2 + w3 * (x_1 * x_2)
```


where $$w_0$$ = _______

Note: We are back to the original data where IQ is in the original scale. (IQ score 100 is recorded as X2=100 and score 150 is recorded as X2=150.)


##Question 11 to 14
Based on the LS coefficients given in  Question 4, given IQ and GPA, the prediction for the salary of a  male student can be written as the following linear combination

$$
m_0 + m_1*\text{GPA} + m_2*\text{IQ} + m_3*(\text{GPA}*\text{IQ})
$$



```{r}
x_1 = 4.0
x_2 = 100
x_3 = 0

yhat_10 = b_0_int + b_1_gpa * x_1 + b_2_iq * x_2 + b_3_gender * x_3 + b_4_gpa_v_iq * x_1 * x_2 + b_5_gpa_v_gender* x_1 * x_3
yhat_10
```

```{r}
(m0 = b_0_int)
(m1 = b_1_gpa)
(m2 = b_2_iq)
(m3 = b_4_gpa_v_iq)

m0 + m1 * x_1 + m2 * x_2 + m3 * (x_1 * x_2)
```


##Question 15-20

Question 15
Suppose we code X_3  = Gender as 1 for Male and 0 for Female.

 
Use the same data to fit a linear regression model. 
```{r}
a_0_int = 85
a_1_gpa = 10
a_2_iq = b_2_iq
a_3_gender = (-35)
a_4_gpa_v_iq = b_4_gpa_v_iq
a_5_gpa_v_gender = 10
```


```{r}
x_1 = 4.0
x_2 = 100
x_3 = 1

a_0_int + (a_1_gpa * x_1) + (a_2_iq * x_2) + (a_3_gender * x_3) + a_4_gpa_v_iq * x_1 * x_2 + a_5_gpa_v_gender* x_1 * x_3
b_0_int + b_1_gpa * x_1 + b_2_iq * x_2 + b_3_gender * x_3 + b_4_gpa_v_iq * x_1 * x_2 + b_5_gpa_v_gender* x_1 * x_3
```

