---
title: "Quiz1"
author: "Brendan Shea"
date: "2022-08-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


For the numerical type of questions, numerical answers should be integers or decimals. Do NOT enter fractions.

If the answer is 0.1, “.1” will still be accepted, although we encourage you to enter “0.1” as the full answer.

Numerical answers should be exact, unless students are asked to round to a certain degree of accuracy.


1. Question 1
Consider tossing a fair six-sided die once. Define event A = {The value on the die is an even number}. What is P(A)?  

```{r}
3/6
```

2. Continue with Question 1. 

Define event C = {The value on the die is 1, 2, 3, or 4}. What’s P(A|C)?  

```{r}
2/4
```

3. Continue with Questions 1-2. 

 Are Event A and Event C independent?  
 
TRUE


4. Question 4
There is a new diagnostic test for a disease that occurs in about 1% of the population. The test is not perfect: people with the disease will be tested positive 99% of the time; however, people without the disease will also be tested positive 3% of the time. Suppose a person, who can be regarded as a random sample from the population, has a positive test result. Then what is the conditional probability that the person has the disease? 

```{r}
prior_pos = 0.01
prior_neg = 1 - prior_pos
sensitivity = 0.99
false_positive = 0.03

pos_term = (prior_pos * sensitivity)
neg_term = (prior_neg * false_positive)

pos_term / (pos_term + neg_term)
```

5. The off-diagonal elements in a variance-covariance matrix contain: 
Covariances (between different variables). The diagnols are the variances.


6. Suppose that the number of hours a household personal computer is used for entertainment per day is normally distributed with mean = two hours and the standard deviation = half an hour.  What is the probability that a household personal computer is used for entertainment between 1.5 and 2.5 hours per day? 

Round your answer to two decimal places.

```{r}
pnorm(1.5, mean = 2, sd = 0.5)
pnorm(2.5, mean = 2, sd = 0.5)

round(
  pnorm(2.5, mean = 2, sd = 0.5) -
  pnorm(1.5, mean = 2, sd = 0.5),
  digits = 2)
```

7. In which of the following situations is the Central Limit Theorem not applicable?  

When the sample size is small and the population is not normal  


8. Let A denote a matrix....What is trace of A * A?
```{r}
A = cbind(
  c(1,2,0),
  c(3,1,2),
  c(1,2,0)
)

A_dotted = A %*% A
sum(diag(A_dotted))

 
```


9. WRONG Consider two vectors. What is the absolute value of the sine between the two vectors?

```{r}
u = c(1,0,2)
v = c(0,2,0)

(d = norm(as.matrix(u), type = "2") * norm(as.matrix(v), type = "2"))

n = u %*% v
sin(acos(n/d)) # This line is changed!



```


Question 10
Consider a function: 
$$
f(x) = (1-x)^2
$$ 

What is the derivative of f evaluated at x=0.5?

```{r}
f = expression((1-x)^2)
x = 0.5
eval(expr= D(f, 'x'))
```

Question 11
Let A be a matrix. What is the (2,2)-th entry of the inverse of A? Round your answer to two decimal places.
```{r}
A = cbind(
  c(4,3),
  c(6,9)
)

(A_inv = solve(A))
round(A_inv[2,2],2)
```

12.
Question 12
  Which two of the following are classification problems?  

1 point

YES We are considering launching a new product and wish to know whether it will be a success or a
failure. We collect data on 20 similar products that were previously launched. For each product
we have recorded whether it was a success or failure, price charged for the product, marketing
budget, competition price, and ten other variables.   

YES We want to feed historical weather data to a learning algorithm to predict tomorrow’s weather:
Sunny, Cloudy, or Rainy.  


    We are interested in predicting the % change in the US dollar in relation to the weekly changes in
the world stock markets. Hence we collect weekly data for all of 2012. For each week we record
the % change in the dollar, the % change in the US market, the % change in the British market,
and the % change in the German market.


  We collect a set of data on the top 500 firms in the US. For each firm we record profit, number of
employees, industry and the CEO salary. We are interested in understanding which factors affect
CEO salar



13. The table below provides a training data set containing six observations, three predictors, and one qualitative response variable.
ANSWER: Rainy for both.

```{r}
predictors = data.frame(
  x1 = c(1,1,3,0,-2,1),
  x2 = c(2,0,1,2,0,0),
  x3 = c(1,0,0,3,1,1)
)

  y = c("S","R","R","R","S","S")
  
test_13 = data.frame(
  x1 = 0,
  x2 = 0,
  x3 = 0
)

knn(train = predictors, test = data.frame(x1=0,x2=0,x3=0), cl = y, k = 1)
knn(train = predictors, test = data.frame(x1=0,x2=0,x3=0), cl = y, k = 3)
```



14.
Question 14
  Which two of the following problems are best addressed using a supervised learning algorithm?  


  Examine a large collection of emails that are known to be spam email, to discover if there are
sub-types of spam email.  


YES: Given 50 articles written by male authors and 50 articles by female authors, learn to predict the
gender of the author of a new manuscript.  


  Given a large dataset of medical records from patients suffering from lung cancer, try to learn
whether there might be different clusters of such patients for which we might tailor separate
treatments.  


YES: Examine the statistics of two football teams (such as historical data of teams’ wins/losses) and
learn to predict which team will win tomorrow’s match.  

15.
Question 15
Consider the data set and weights given below

$$
x = (0.82, 0.24, -0.30, -0.36), 
w = (2, 1, 3, 1)
$$

What's the value of u that minimizes the following equation

$$\sum_{i=1}^4 w_i (x_i - u)^2$$

```{r}
x = c(0.82, 0.24, -0.30, -0.36)
w = c(2,1,3,1)

f_15 = function(u, w, x){
  sum = 0
  for(i in 1:4){
    sum = sum +
      w[i] * (x[i] - u)^2
  }
  return(sum)
}

optimize(f_15, x = x, w = w, lower = -100, upper = 1000)
round(31/350,2)
```



Question 16
Let A1, A2A1,A2 denote two events. What can we say about the chance both occur:

```{r}
0.95 * 0.9 ## worst case is independent

```

We know it is greater than 0.85 (and nothing else)